A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2023-08-09 tidyindex.qmd
pipeline paragraph of section 2

For constructing indexes, rather than implementing each index separately, as it has been done for some indexes, the pipeline approach introduces explicit and standalone modules for analysts to assemble indexes. Analysts can then rearrange the order of modules in the pipeline or insert new modules to experiment with possible new index definition or to understand index behaviour. Also, the readability of the pipeline code can complement conventional diagrams and textual information to understand the how indexes are constructed. The uniform interface of pipeline allows analysts to assemble a diverse array of indexes with a shared set of components, reducing the learning curve for adopting new implementation for different indexes.

@@

# Details of the index pipline {#sec-index-pipeline}

@@

```{r}
library(sf)
qld_lga <- absmapsdata::lga2018 %>%
  filter(state_name_2016 == "Queensland") %>%
  rmapshaper::ms_simplify(keep = 0.3) %>%
  mutate(lga_name_2018 = str_replace(lga_name_2018, " \\([S|R|C|]\\)", ""),
         lga_name_2018 = str_replace(lga_name_2018, " \\(Qld\\)", ""))
```

>>>>

For constructing indexes, the pipeline approach adopts explicit and standalone modules that can be assembled in different ways. Index developers can choose the appropriate modules and arrange them accordingly to generate the data pipeline that is needed for their purpose. The pipeline approach provides many advantages:

- makes the computation more transparent, and thus more easily debugged.
- allows for rapidly processing new data to check how different features, like outliers, might affect the index value.
- provides the capacity to measure uncertainty by computing confidence intervals from multiple samples as generated by bootstrapping to original data.
- enables systematic comparison of surrogate indexes designed to measure the same phenomenon.
- it may even be possible to automate diagramatic explanations and documentation of the index.

Adoption of this pipeline approach would provide uniformity to the field of index development, research and application.

<!-- They might even  insert new modules to experiment with possible new index definitions or to understand an index's behaviour. Also, the readability of the pipeline code can complement conventional diagrams and textual information to understand the how indexes are constructed. The uniform interface of pipeline allows analysts to assemble a diverse array of indexes with a shared set of components, reducing the learning curve for adopting new implementation for different indexes.-->

@@

# Details of the index pipeline {#sec-index-pipeline}

@@

```{r}
library(sf)
# options(timeout = 1000)
# remotes::install_github("wfmackey/absmapsdata")
qld_lga <- absmapsdata::lga2018 %>%
  filter(state_name_2016 == "Queensland") %>%
  rmapshaper::ms_simplify(keep = 0.3) %>%
  mutate(lga_name_2018 = str_replace(lga_name_2018, " \\([S|R|C|]\\)", ""),
         lga_name_2018 = str_replace(lga_name_2018, " \\(Qld\\)", ""))
```

### 2023-08-09 tidyindex.tex
pipeline paragraph of section 2

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{orcidlink}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Ursula Laa; Nicolas Langrené; Patricia Menéndez},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, interior hidden, breakable, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

For constructing indexes, rather than implementing each index
separately, as it has been done for some indexes, the pipeline approach
introduces explicit and standalone modules for analysts to assemble
indexes. Analysts can then rearrange the order of modules in the
pipeline or insert new modules to experiment with possible new index
definition or to understand index behaviour. Also, the readability of
the pipeline code can complement conventional diagrams and textual
information to understand the how indexes are constructed. The uniform
interface of pipeline allows analysts to assemble a diverse array of
indexes with a shared set of components, reducing the learning curve for
adopting new implementation for different indexes.

\hypertarget{sec-index-pipeline}{%
\section{Details of the index pipline}\label{sec-index-pipeline}}

@@

\hypertarget{tbl-gggi-weights}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.5455}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1299}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1169}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1169}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0909}}@{}}
\caption{\label{tbl-gggi-weights}Weights of the fourteen variables in
Global Gender Gap Index}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
V-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
D-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Weight
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
V-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
D-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Weight
\end{minipage} \\
\midrule()
\endhead
Labour force participation & Economy & 0.199 & 0.25 & 0.050 \\
Wage equality for similar work & Economy & 0.310 & 0.25 & 0.078 \\
Estimated earned income & Economy & 0.221 & 0.25 & 0.055 \\
Legislators senior officials and managers & Economy & 0.149 & 0.25 &
0.037 \\
Professional and technical workers & Economy & 0.121 & 0.25 & 0.030 \\
Literacy rate & Education & 0.191 & 0.25 & 0.048 \\
Enrolment in primary education & Education & 0.459 & 0.25 & 0.115 \\
Enrolment in secondary education & Education & 0.230 & 0.25 & 0.058 \\
Enrolment in tertiary education & Education & 0.121 & 0.25 & 0.030 \\
Sex ratio at birth & Health & 0.693 & 0.25 & 0.173 \\
Healthy life expectancy & Health & 0.307 & 0.25 & 0.077 \\
Women in parliament & Politics & 0.310 & 0.25 & 0.078 \\
Women in ministerial positions & Politics & 0.247 & 0.25 & 0.062 \\
Years with female head of state & Politics & 0.443 & 0.25 & 0.111 \\
\bottomrule()
\end{longtable}

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{orcidlink}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Ursula Laa; Nicolas Langrené; Patricia Menéndez},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

For constructing indexes, the pipeline approach adopts explicit and
standalone modules that can be assembled in different ways. Index
developers can choose the appropriate modules and arrange them
accordingly to generate the data pipeline that is needed for their
purpose. The pipeline approach provides many advantages:

\begin{itemize}
\tightlist
\item
  makes the computation more transparent, and thus more easily debugged.
\item
  allows for rapidly processing new data to check how different
  features, like outliers, might affect the index value.
\item
  provides the capacity to measure uncertainty by computing confidence
  intervals from multiple samples as generated by bootstrapping to
  original data.
\item
  enables systematic comparison of surrogate indexes designed to measure
  the same phenomenon.
\item
  it may even be possible to automate diagramatic explanations and
  documentation of the index.
\end{itemize}

Adoption of this pipeline approach would provide uniformity to the field
of index development, research and application.

\hypertarget{sec-index-pipeline}{%
\section{Details of the index pipeline}\label{sec-index-pipeline}}

@@

\hypertarget{tbl-gggi-weights}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.5455}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1299}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1169}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1169}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0909}}@{}}
\caption{\label{tbl-gggi-weights}Weights of the fourteen variables in
Global Gender Gap Index}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
V-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
D-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Weight
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
V-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
D-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Weight
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Labour force participation & Economy & 0.199 & 0.25 & 0.050 \\
Wage equality for similar work & Economy & 0.310 & 0.25 & 0.078 \\
Estimated earned income & Economy & 0.221 & 0.25 & 0.055 \\
Legislators senior officials and managers & Economy & 0.149 & 0.25 &
0.037 \\
Professional and technical workers & Economy & 0.121 & 0.25 & 0.030 \\
Literacy rate & Education & 0.191 & 0.25 & 0.048 \\
Enrolment in primary education & Education & 0.459 & 0.25 & 0.115 \\
Enrolment in secondary education & Education & 0.230 & 0.25 & 0.058 \\
Enrolment in tertiary education & Education & 0.121 & 0.25 & 0.030 \\
Sex ratio at birth & Health & 0.693 & 0.25 & 0.173 \\
Healthy life expectancy & Health & 0.307 & 0.25 & 0.077 \\
Women in parliament & Politics & 0.310 & 0.25 & 0.078 \\
Women in ministerial positions & Politics & 0.247 & 0.25 & 0.062 \\
Years with female head of state & Politics & 0.443 & 0.25 & 0.111 \\
\end{longtable}
